---
title: "Many Local Maxima"
author: "Guanting Wei"
date: "Sep.25.2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Formula
Density function:
\[
\begin{aligned}
f(x;\theta)=\frac{1-\cos(x-\theta)}{2\pi},
\\0\leq x\leq 2\pi,\theta\in (-\pi,\pi)
\end{aligned}
\eqno{(1.1)}
\]

The likelihood funcion:
\[
\begin{aligned}
L(\theta)=\prod_{i=1}^{n}f(X_{i};\theta)
\\0\leq x\leq 2\pi,\theta\in (-\pi,\pi)
\end{aligned}
\eqno{(1.2)}
\]

The loglikelihood funcion:
\[
\begin{aligned}
{l(\theta)}=\ln L(\theta)=\sum_{i=1}^{n}\ln f(X_{i};\theta)=\sum_{i=1}^{n}\ln[\frac{1-\cos(x-\theta)}{2\pi}]
\\0\leq x\leq 2\pi,\theta\in (-\pi,\pi)
\end{aligned}
\eqno{(1.3)}
\]

Compute the differential of loglikelihood funcion:

First derivative:
\[
\begin{aligned}
{l'(\theta)}=\sum_{i=1}^{n}\frac{\sin(X_{i}-\theta)}{1-\cos(X_{i}-\theta)}
\\0\leq x\leq 2\pi,\theta\in (-\pi,\pi)
\end{aligned}
\eqno{(1.4)}
\]

Second derivative:
\[
\begin{aligned}
{l''(\theta)}=\sum_{i=1}^{n}\frac{1}{[1-\cos(X_{i}-\theta)]^2}\
\\0\leq x\leq 2\pi,\theta\in (-\pi,\pi)
\end{aligned}
\eqno{(1.5)}
\]

# 2. Plot The Log-likelihood Function

Sample:
```{r}
X <- c(3.91, 4.85, 2.28, 4.06, 3.70, 4.04, 5.46, 3.53, 2.28, 1.96,
       2.53, 3.88, 2.22, 3.47, 4.82, 2.46, 2.99, 2.54, 0.52)
```

Define loglikelihood function and its differential function
```{r}
f=function(x,theta)(1-cos(x-theta))/(2*pi)

L=function(x,theta){
  prod=1;
  for (i in 1:length(x)){
    prod = prod*((1-cos(x[i]-theta))/(2*pi));
  }
  prod
}

l=function(x,theta){log(L(x,theta))}

l1=function(x,theta){
  sum=0;
  for(i in 1:length(x)){
    sum=sum+sin(x[i]-theta)/(1-cos(x[i]-theta))
  }
  sum
}

l2=function(x,theta){
  sum=0;
  for(i in 1:length(x)){
    sum=sum+1/(1-cos(x[i]-theta))^2
  }
  sum
}

```
## plot
```{r}
library("ggplot2")
ggplot(data.frame(x=c(-pi,pi)),aes(x=x)) + 
  stat_function(fun=function(theta) l(X,theta)) + 
  labs(x=expression("theta"),y="l")
```

# 3. Method-of-Moments

\[
\begin{aligned}
E(X|\theta)&=\int_{0}^{2\pi}xf(x;\theta)\mathrm{d}x\\&=\int_{0}^{2\pi}x\frac{1-\cos(x-\theta)}{2\pi}\mathrm{d}x\\&=\frac{1}{2\pi}\int_{0}^{2\pi}x\mathrm{d}[x-\sin (x-\theta)]\\&=\frac{1}{2\pi}\left \{ x[x-\sin (x-\theta)]\bigg|_{0}^{2\pi}-\int_{0}^{2\pi}x-\sin (x-\theta)\mathrm{d}x \right \}\\&=\pi+\sin (\theta)
\end{aligned}
\]
```{r}
E=function(theta){pi+sin(theta)}
X_n=mean(X)
theta_n=c(0,0)
theta_n[1]=asin(mean(X)-pi)
theta_n[2]=pi-theta_n[1]
theta_n[1]
theta_n[2]
```
So we can get$\tilde{\theta_n}=0.09539407\ or\ 3.046199$

# 4. MLE
```{r}
r1=c(0,0)
count1=matrix(0,1, length(theta_n))
for(i in 1:length(theta_n)) {
  r1[i]=theta_n[i]
  while (abs(l1(X,r1[i]))>.Machine$double.eps&&count1[i]<10000) {
    temp=r1[i]-l1(X,r1[i])/l2(X,r1[i])
    r1[i]=temp
    count1[i]=count1[i]+1
  }
}
table1=rbind(theta_n,r1)
rownames(table1)=c('theta_0',"root")
library(pander)
set.caption("theta_0 and roots")
pander(table1)
```


# 5. Different Initial $\theta$
```{r}
theta_0=c(-2.7,2.7)
r2=c(0,0)
count2=matrix(0,1, length(theta_0))
for(i in 1:length(theta_0)) {
  r2[i]=theta_0[i]
  while (abs(l1(X,r2[i]))>0.00000001&&count2[i]<10000) {
    temp=r2[i]-l1(X,r2[i])/l2(X,r2[i])
    r2[i]=temp
    count2[i]=count2[i]+1
  }
}
table2=rbind(theta_0,r2)
rownames(table2)=c('theta_0',"root")
library(pander)
set.caption("theta_0 and roots")
pander(table2)
```

# 6. Initial $\theta$ from $-\pi$ to $\pi$
```{r}
i=seq(-pi, pi, length=200)
r=matrix(0,1, length(i))
count=matrix(0,1, length(i))
for(k in 1:length(i)) {
  r[k]=i[k]
  while (abs(l1(X,r[k]))>0.00000001&&count[k]<10000) {
    temp=r[k]-l1(X,r[k])/l2(X,r[k])
    r[k]=temp
    count[k]=count[k]+1
  }
}
library(pander)
table3=rbind(i,r)
rownames(table3)=c("i","root")
set.caption("200 equally spaced initial thetas from -pi to pi")
pander(table3)
plot(i,r,xlab = NULL, ylab = "root",main="200 equally spaced initial thetas from -pi to pi")
```

Count the roots
```{r}
r_simp=round(r,4)
pander(table(r_simp))
```

So the partion can be
[1:11] [12] [13] [14:18] [19] [20:24] [25] [26:29] [30] [31:53] [54]
[55]  [56:74]  [75:116]  [117]  [118  162]  [163]  [164:170]  [171]  [172]
[173]  [174:178]  [179]  [180]  [181]  [182:194]  [195]  [196:200]
