---
title: "Modeling beetle data"
author: "Guanting Wei"
date: "Sep.26.2018"
output: pdf_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Squared errors
```{r}
beetles <- data.frame(
    days    = c(0,  8,  28,  41,  63,  69,   97, 117,  135,  154),
    beetles = c(2, 47, 192, 256, 768, 896, 1120, 896, 1184, 1024))
```

\[
N_t=f(t)=\frac{KN_0}{N_0+(K-N_0)e^{-rt}}
\]

According to the data, $N_0=2$, then
\[
N_t=f(t)=\frac{2K}{2+(K-2)e^{-rt}}
\]

Squared Errors:
\[
SE=\sum [N_t-\frac{2K}{2+(K-2)e^{-rt}}]^2
\]

In order solve the problem, we need to find range of K and r. We already know K is parameter that represents the population carrying capacity of the environment, so K should be bigger than$N_t$.Then I choose K ranging from 1300 to 1600 and then we decide r by using $N_t$,t and K

\[
r=\frac{1}{t}\ln \frac{N_t(K-2)}{2(K-N_t)}
\]

```{r}
t=beetles$days
n=beetles$beetles
r=function(t,k,n){
  r=1/t*log(n*(k-2)/(2*(k-n)))
}
k=1300:1600
a=matrix(0,length(k),10)
for(m in 1:length(k)){
  for(i in 2:10){
  a[m,i]=r(t[i],k[m],n[i])
  }
}
max=max(a[a>0])
min=min(a[a>0])
max
min
```

Then we choose t from 0.04 to 0.4. However, $N_t$ decreases after 97 days, so the environment is overloaded. As a result, we change K to range from 600 to 1600

```{r}
k=600:1600
r=seq(0.04,0.4,length=1000)
```

Plot the contour plot depending on K and r above:
```{r}
se=function(k,r){
  se=0
  for(i in 1:length(beetles$days)){
    se=se+(n[i]-(2*k)/(2+(k-2)*exp(-r*t[i])))^2
  }
  se
}
b=matrix(0,length(k),length(r))
for(m in 1:length(k)){
  for(i in 1:length(r)){
    b[m,i]=se(k[m],r[i])
  }
}
contour(k,r,b,xlab="K",ylab="r",main="contour plot of squared errors")
```

# 2. MLE 
We assume that$\log N_t\sim N(\log f(t),\sigma ^2)$
The Likelihood function is:
\[
L(k,r,\sigma^2)=\prod_{i=1}^{n}\frac{1}{\sqrt{2\pi\sigma^2}}e^{-\frac{(\log N_i-\log\frac{2K}{2+(K-2)e^{-rt}})^2}{2\sigma^2}}
\]
The Loglikelihood function is:
\[
l(k,r,\sigma^2)=-\frac{n}{2}\log (2\pi\sigma^2)-\sum_{i=1}^{n}{\frac{(\log N_i-\log\frac{2K}{2+(K-2)e^{-rt}})^2}{2\sigma^2}}
\]

```{r}
m=function(x){
  k=x[1]
  r=x[2]
  sigma_2=x[3]
  sum=0
  for(i in 1:10){
  sum=sum+(-(log(2*pi*sigma_2))/2-(log(n[i])-log((2*k)/(2+(k-2)*exp(-r*t[i]))))^2/(2*sigma_2))
  }
  -sum
}
options(warn=-1)
opt=optim(c(800,0.2,4),m,method="BFGS",hessian=TRUE)
opt
k0=opt$par[1]
r0=opt$par[2]
sigma0_2=opt$par[3]
var=diag(solve(opt$hessian))
```
$\theta=(r,K,\sigma^2)=(0.1958194,800,0.4151383)\\$
$var(k)=6.016225e+04\\$
$var(r)=4.777223e-03\\$
$var(\sigma^2)=3.451498e-02\\$
